---
title: "SLF_SCP_LLTemp_analysis"
author: "Clarissa Moore"
date: "2024-11-22"
output: html_document
---
 
## load data

```{r}
rm(list=ls()) # clear environment window
SCP <- read.csv("data/SLF_SCP_data.csv")
LLTemp <- read.csv("data/SLF_LLTemp_BVT_data.csv")
hatch <- read.csv("data/SLF_LLTemp_hatch_data.csv")
```

## load packages

```{r}
#this will download any packages if you don't already have them
packages <- c("tidyverse", "betareg", "survival", "lmerTest", "emmeans", "multcompView", "glmmTMB", "pwr", "ggh4x", "multcomp", "flextable", "car", "MASS", 'dunn.test', "dplyr","rstatix", "viridis")

for (package in packages) {
  if (!requireNamespace(package, quietly = TRUE)) {
    install.packages(package)
  }
}
#load them
lapply(packages, library, character.only = TRUE)

#make color pallette
set.seed(123)  # for reproducibility
# 10 evenly spaced colors from the viridis palette
cols <- viridis(10)
# randomize their order
Palette <- sample(cols)
Palette

```

## clean data
```{r}
## bvt
LLTemp <- LLTemp[is.na(LLTemp$remove),] #remove dried eggs\

LLTemp <- LLTemp[!(LLTemp$treatment %in% c("Live Control", "Dead Control")), ] #remove controls, since these were only used to calculate viability, done in Excel

LLTemp <- LLTemp %>% #add right and left bounds of interval censoring; 
  mutate(left = ifelse(viability == 1, temp, Inf),
         right = ifelse(viability == 1, -Inf, temp)) #if live, left = removal temp, right = -infinity, if dead, left = infinity, right = removal temp

LLTemp$sitemonth <- paste(LLTemp$site, LLTemp$month, sep = "") #create a combined site and month variable

## hatch
hatch$sitemonth <- paste(hatch$site, hatch$month, sep = "") #create a combined site and month variable

prop_hatch <- hatch #save version of data before making it binary

hatch <- hatch %>% #make hatch data binary
  rowwise() %>%
  mutate(hatched = list(c(rep(1, num_hatched), rep(0, num_unhatched)))) %>%
  unnest(hatched) %>%
  ungroup()

hatch <- hatch %>% #add right and left bounds of interval censoring; 
  mutate(left = ifelse(hatched == 1, temp, Inf),
         right = ifelse(hatched == 1, -Inf, temp)) #if live, left = removal temp, right = -infinity, if dead, left = infinity, right = removal temp

prop_hatch <- prop_hatch[!(prop_hatch$month == "December" & prop_hatch$run %in% c("third", "fourth")), ] 
hatch <- hatch[!(hatch$month == "December" & hatch$run %in% c("third", "fourth")), ] #remove the additional controls that were nto assigned a spsecific run

## SCP
SCP$sitemonth <- paste(SCP$site, SCP$month, sep = "") #create a combined site and month variable

```

# BVT LLTEMP

## raw data viz

```{r}
PropData <- LLTemp %>%
  group_by(site, month, temp, run)%>%
  dplyr::summarize(propalive = sum(viability) / n(), .groups = 'drop')

ggplot(PropData, aes(x = temp, y = propalive, color= run)) +
  geom_point() +
  geom_line() +
  facet_grid(site ~ month) +
  scale_x_reverse()

```

## Kaplan-Meier

```{r}
#generate Kaplan-Meier curves
LLTemp<-LLTemp[!(LLTemp$month == "February" & LLTemp$run == "first"),]
surv_obj <- with(LLTemp, Surv(time = right, time2 = left, type = "interval2"))
head(surv_obj, 100)
km_fit <- survfit(surv_obj ~ sitemonth + run, data = LLTemp)
km_fit ##can see LT50

#exract data
survd <- data.frame(temp = km_fit$time,
                        propalive = km_fit$surv, 
                        lower = km_fit$lower, 
                        upper = km_fit$upper)
survd_newdata <- data.frame(strata = km_fit$strata)
survd_newdata <- survd_newdata %>%
  rownames_to_column(var = "var") %>%
  separate(col = "var", 
           into = c("sitemonth", "run"),
           sep = ",") %>%
  mutate(sitemonth = sub(".*=", "", sitemonth)) %>%
  mutate(run = sub(".*=", "", run)) %>%
  slice(rep(row_number(), strata))
survd <- cbind(survd, survd_newdata[,-3]) %>%
  mutate(
    sitemonth = str_replace_all(sitemonth, "\\s+", ""),
    run = str_replace_all(run, "\\s+", ""),
  propalive = 1-propalive, #take the inverse, as to display survival not mortality
  lower = 1-lower, #take the inverse
  upper = 1-upper) #take the inverse

ggplot()+
  geom_point(data = survd, aes(x = temp, y = propalive, color = run))+
  geom_line(data = survd, aes(x = temp, y = propalive, color = run)) +
  geom_ribbon(data = survd, aes(x = temp, ymin = lower, ymax = upper, fill = run), alpha = 0.1)+
  facet_wrap(sitemonth~.)+
  scale_x_reverse() 

survd <- survd %>% mutate(propalive = ifelse(propalive == 0, 0.001, ifelse(propalive == 1, 0.999, survd$propalive)))

```

## BVT-LLTemp model on JUST 2026 data

```{r}
survd2026<-survd[survd$sitemonth %in% c("CTJanuary", "CTMarch"),]
unique(survd2026$sitemonth)

LLTempModel <- glmmTMB(propalive ~ temp * sitemonth + (1|run), data = survd2026, family = beta_family(link = "logit"))

LLTempModel <- glmmTMB(propalive ~ temp + (1+temp|sitemonth), data = survd2026, family = beta_family(link = "logit"))

LLTempModel <- glmmTMB(propalive ~ temp + (1|sitemonth) + (0+temp|sitemonth), data = survd2026, family = beta_family(link = "logit"))
#scale temperature
#cox proportional hazard model

LLTempModel <- glmmTMB(propalive ~ temp + (1|sitemonth), data = survd2026, family = beta_family(link = "logit"))

summary(glht(LLTempModel), adjust = "bonferroni")
summary(LLTempModel)
```






## BVT-LLTemp model on JUST Connecticut data to see effect of year

```{r}
survdCT<-survd[survd$sitemonth %in% c("CTJanuary", "CTMarch", "MilfordJanuary", "Milford February"),]

LLTempModel <- glmmTMB(propalive ~ temp * sitemonth + (1|run), data = survdCT, family = beta_family(link = "logit"))
summary(glht(LLTempModel), adjust = "bonferroni")
```

## beta reg with month for all data in combo

```{r}

LLTempModel <- glmmTMB(propalive ~ temp * sitemonth + (1|run), data = survd, family = beta_family(link = "logit"))
summary(glht(LLTempModel), adjust = "bonferroni")



# Estimated marginal means for sitemonth at temp = 0
emm <- emmeans(LLTempModel, ~ sitemonth, at = list(temp = 0))



#see slopes for each sitemonth
emtrends(LLTempModel, ~ sitemonth, var = "temp")

# Get comparisons slopes of temp between  sitemonths
slopes <- emtrends(LLTempModel, ~ sitemonth, var = "temp")
pairs(slopes)

# All pairwise comparisons with Bonferroni correction
summary(slopes, infer = TRUE)

#sample sizes
table(LLTemp$sitemonth)
table(survd$sitemonth)

#calculate proportion variance explained by random effects
var <- var(residuals(LLTempModel))
randovar <- 3.013e-08

randovar/(randovar+var)
#4.300027e-07 or 0.00000043 or 0.000043%

#find the dispersion parameter theta/phi
sigma(LLTempModel)

```

## plot and quantiles with month

```{r}
#make predictions
LLTempModel <- glmmTMB(propalive ~ temp * sitemonth + (1|run), data = survd, family = beta_family(link = "logit"))
newdata <- expand.grid(sitemonth = unique(survd$sitemonth), temp = c(seq(10, -1, by = -1), seq(-1, -75, by = -0.01))) 

#to acheieve confidence interval between 0 and 1, calculate fit and SE with predict on the LINK function (to get logit link values), then back tranform them to true values
LLTempP <- cbind(newdata,
                 predict(LLTempModel, newdata, type = "link", se.fit = T, re.form = NA))
LLTempP$lower <- LLTempP$fit - 1.96 * LLTempP$se.fit
LLTempP$upper <- LLTempP$fit + 1.96 * LLTempP$se.fit
LLTempP$fit <- plogis(LLTempP$fit)
LLTempP$upper <- plogis(LLTempP$upper)
LLTempP$lower <- plogis(LLTempP$lower)

#retrieve quantiles
calculate_quantiles <- function(data) {
  list(
    data.frame(
      quantile = "50%", 
      estimate = data[which.min(abs(data$fit - 0.5)), ]$temp,
      lower = data[which.min(abs(data$lower - 0.5)), ]$temp, 
      upper = data[which.min(abs(data$upper - 0.5)), ]$temp
    ),
    data.frame(
      quantile = "90%", 
      estimate = data[which.min(abs(data$fit - 0.1)), ]$temp,
      lower = data[which.min(abs(data$lower - 0.1)), ]$temp, 
      upper = data[which.min(abs(data$upper - 0.1)), ]$temp
    )
  )
}
quantiles <- LLTempP %>%
  group_by(sitemonth) %>%
  do(bind_rows(calculate_quantiles(.))) %>%
  ungroup()
quantiles[quantiles$quantile == "90%",]

#save quantiles as a data file 

# write.csv(quantiles, file ="C:/Users/venet001/Documents/Entomology/Venette/SLF/Outputs_from_R_Clarissa/SLF_LLTemp_BVT_quantiles.csv") 
#Clarissa's original 
write.csv(quantiles, file = "C:/Users/CLAHM/Downloads/SLF/Data/SLF_LLTemp_BVT_quantiles.csv")

# #trying to get "real" 95%CI
# LLTempP <- LLTempP %>%
#   mutate(Variance = (fit*(1- fit))/(1+1.04)) %>%
#   mutate(lower2 = fit - (1.96 * (sqrt(Variance) / sqrt(74))), upper2 = fit + (1.96 * (sqrt(Variance) / sqrt(74))))

#plot
LLTempP <- LLTempP %>%
  mutate(lower2 = ifelse(lower<0,0,lower),
         upper2 = ifelse(upper>1,1,upper))
LLTempP$sitemonth <- factor(LLTempP$sitemonth, levels = c("MilfordDecember","MilfordJanuary","CTJanuary", "CTMarch", "MilfordFebruary","WinchesterDecember","WinchesterJanuary","RoanokeDecember","RoanokeJanuary"))
survd$sitemonth <- factor(survd$sitemonth, levels = c("MilfordDecember","MilfordJanuary","CTJanuary", "CTMarch","MilfordFebruary","WinchesterDecember","WinchesterJanuary","RoanokeDecember","RoanokeJanuary"))

ggplot() +
  geom_point(data = survd, aes(x = temp, y = propalive, color = run)) +
  geom_line(data = LLTempP, aes(x = temp, y = fit)) +
  geom_ribbon(data = LLTempP, aes(x = temp, ymin = lower2, ymax = upper2), alpha = 0.1)+
  facet_wrap2(sitemonth~., nrow=2, axes = "all", remove_labels = "all")+
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line(),
    legend.position = c(.85, 0.25)) +
  labs(color = "", x = "Temperature (°C)", y = "Proportion Alive")+
  scale_x_reverse()

#overlayed plot
ggplot() +
  geom_point(data = survd, aes(x = temp, y = propalive, color = sitemonth)) +
  geom_line(data = LLTempP, aes(x = temp, y = fit, color = sitemonth)) +
  geom_ribbon(data = LLTempP, aes(x = temp, ymin = lower2, ymax = upper2, fill = sitemonth), alpha = 0.1)+
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line()) +
  labs(color = "", fill = "", x = "Temperature (°C)", y = "Proportion Alive")+
  scale_x_reverse(limits=c(0,-60))+
  scale_y_continuous(breaks = c(1, 0.9, 0.75, 0.5, 0.25, 0), labels = c("0", "0.1","0.25","0.5","0.75","1"))+
  scale_fill_manual(values = Palette)+
  scale_color_manual(values = Palette)

    ```

## beta reg without month

```{r}
#make model
survd$run2 <- paste(survd$sitemonth, survd$run, sep = "") #when pooling we need to change run from "first", "second" to unique, so gluing run to sitemonth
LLTempModel_pool <- glmmTMB(propalive ~ temp + (1|run2), data = survd, family = beta_family(link = "logit"))
summary(LLTempModel_pool)

#compare AICs
AIC(LLTempModel, LLTempModel_pool)

#calculate proportion variance explained by random effects
var <- var(residuals(LLTempModel_pool))
randovar <- 9.618e-10 #the Variance next to run2 in model output
var
randovar/(randovar+var)
#1.332563e-08 or 0.00000001332563 or 0.00000133

#find the dispersion parameter theta/phi
sigma(LLTempModel_pool)

```

## plot and quantiles without month

```{r}
#make predictions
newdata <- data.frame(temp = c(seq(10, -12, by = -1), seq(-12, -40, by = -0.01), seq(-40, -55, by = -1))) 
LLTempP_pool <- cbind(newdata,
                 predict(LLTempModel_pool, newdata, type = "link", se.fit = T, re.form = NA))

LLTempP_pool$lower <- LLTempP_pool$fit - 1.96 * LLTempP_pool$se.fit
LLTempP_pool$upper <- LLTempP_pool$fit + 1.96 * LLTempP_pool$se.fit
LLTempP_pool$fit <- plogis(LLTempP_pool$fit)
LLTempP_pool$upper <- plogis(LLTempP_pool$upper)
LLTempP_pool$lower <- plogis(LLTempP_pool$lower)

#retrieve quantiles
quantiles_pool <- rbind(data.frame(quantile = "50%", 
      estimate = LLTempP_pool[which.min(abs(LLTempP_pool$fit - 0.5)), ]$temp,
      lower = LLTempP_pool[which.min(abs(LLTempP_pool$lower - 0.5)), ]$temp, 
      upper = LLTempP_pool[which.min(abs(LLTempP_pool$upper - 0.5)), ]$temp),
    data.frame(quantile = "90%", 
      estimate = LLTempP_pool[which.min(abs(LLTempP_pool$fit - 0.1)), ]$temp,
      lower = LLTempP_pool[which.min(abs(LLTempP_pool$lower - 0.1)), ]$temp, 
      upper = LLTempP_pool[which.min(abs(LLTempP_pool$upper - 0.1)), ]$temp))
# LLTempP_pool <- LLTempP_pool %>%
#   mutate(lower2 = ifelse(lower<0,0,lower),
#          upper2 = ifelse(upper>1,1,upper))
ggplot() +
  geom_point(data = survd, aes(x = temp, y = propalive)) +
  geom_line(data = LLTempP_pool, aes(x = temp, y = fit)) +
  geom_ribbon(data = LLTempP_pool, aes(x = temp, ymin = lower, ymax = upper), alpha = 0.1)+
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line()) +
  labs(color = "", fill = "", x = "Temperature(°C)", y = "Proportion Alive")+
  scale_x_reverse()

quantiles_pool

#before 2026 data was added:
#LT50 was -27.25 [-28.21, -26.13]	
#LT90 was -35.31 [-37.76, -32.97]

```    

## with Abbott's correction

```{r}
maxs<- survd %>%
  group_by(run2) %>%
  summarise(max= max(propalive))

maxCTJ3 <- max(survd[survd$run2 == "CTJanuarythird",]$propalive)/0.999
#divide by 0.999 because algebraically it makes sens, will make highest value 0.999 like rest of data, not 1

#check
max(survd[survd$run2 == "CTJanuarythird",]$propalive)/maxCTJ3

survd <- survd %>%
  mutate(propalive2 = ifelse(run2 == "CTJanuarythird", propalive/maxCTJ3, propalive))
```

```{r}
#remake model with this propalive2:

#make model
survd$run2 <- paste(survd$sitemonth, survd$run, sep = "") #when pooling we need to change run from "first", "second" to unique, so gluing run to sitemonth
LLTempModel_pool <- glmmTMB(propalive2 ~ temp + (1|run2), data = survd[survd$propalive2 <0,], family = beta_family(link = "logit"))
summary(LLTempModel_pool)

#compare AICs
AIC(LLTempModel, LLTempModel_pool)

#calculate proportion variance explained by random effects
var <- var(residuals(LLTempModel_pool))
randovar <- 9.618e-10 #the Variance next to run2 in model output
var
randovar/(randovar+var)
#1.332563e-08 or 0.00000001332563 or 0.00000133

#find the dispersion parameter theta/phi
sigma(LLTempModel_pool)
```

```{r}

#make predictions
newdata <- data.frame(temp = c(seq(10, -12, by = -1), seq(-12, -40, by = -0.01), seq(-40, -55, by = -1))) 
LLTempP_pool <- cbind(newdata,
                 predict(LLTempModel_pool, newdata, type = "link", se.fit = T, re.form = NA))

LLTempP_pool$lower <- LLTempP_pool$fit - 1.96 * LLTempP_pool$se.fit
LLTempP_pool$upper <- LLTempP_pool$fit + 1.96 * LLTempP_pool$se.fit
LLTempP_pool$fit <- plogis(LLTempP_pool$fit)
LLTempP_pool$upper <- plogis(LLTempP_pool$upper)
LLTempP_pool$lower <- plogis(LLTempP_pool$lower)


#retrieve quantiles
quantiles_pool <- rbind(data.frame(quantile = "50%", 
      estimate = LLTempP_pool[which.min(abs(LLTempP_pool$fit - 0.5)), ]$temp,
      lower = LLTempP_pool[which.min(abs(LLTempP_pool$lower - 0.5)), ]$temp, 
      upper = LLTempP_pool[which.min(abs(LLTempP_pool$upper - 0.5)), ]$temp),
    data.frame(quantile = "90%", 
      estimate = LLTempP_pool[which.min(abs(LLTempP_pool$fit - 0.1)), ]$temp,
      lower = LLTempP_pool[which.min(abs(LLTempP_pool$lower - 0.1)), ]$temp, 
      upper = LLTempP_pool[which.min(abs(LLTempP_pool$upper - 0.1)), ]$temp))
# LLTempP_pool <- LLTempP_pool %>%
#   mutate(lower2 = ifelse(lower<0,0,lower),
#          upper2 = ifelse(upper>1,1,upper))
ggplot() +
  geom_point(data = survd, aes(x = temp, y = propalive2)) +
  geom_line(data = LLTempP_pool, aes(x = temp, y = fit)) +
  geom_ribbon(data = LLTempP_pool, aes(x = temp, ymin = lower, ymax = upper), alpha = 0.1)+
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line()) +
  labs(color = "", fill = "", x = "Temperature(°C)", y = "Proportion Alive")+
  scale_x_reverse()

quantiles_pool #.11 increase in lt50, .19 decrease in lt90

#before 2026 data was added:
#LT50 was -27.25 [-28.21, -26.13]	
#LT90 was -35.31 [-37.76, -32.97]

```    

# HATCH LLTEMP

## data exploration

```{r}

ggplot(prop_hatch, aes(x = temp, y = prop_hatch, color = run))+
  geom_point()+
  geom_line()+
  facet_wrap(.~sitemonth, nrow = 2)+
  scale_color_manual(values = Palette)+
  scale_x_reverse()+
  theme(legend.position = c(.85,.25))

ggplot(prop_hatch, aes(x = sitemonth, y = num_eggs))+
  geom_boxplot()+
  geom_point()

summary(prop_hatch$num_eggs)
#number of eggs per egg mass ranged from 11 to 84, 32.04 on average. Some pieces of bark with unusually high number of eggs likely had multiple egg masses, though they were laid adjacent such that we could not deliniate them confidently and thus considered them one mass.

shapiro.test(prop_hatch[prop_hatch$site == "Milford",]$num_eggs) #see that we should use non-parametric test

kruskal.test(prop_hatch$num_eggs~prop_hatch$site) #run kruskal-wallis test

dunn.test(prop_hatch$num_eggs,prop_hatch$site) #run post-hoc dunn.test
by(prop_hatch$num_eggs,prop_hatch$site, mean)
prop_hatch%>%
  group_by(site) %>%
  get_summary_stats(num_eggs, type = "mean_se") #find group means +- SEs

#The Kruskal-Wallis test for the effect of site on egg mass size (H = 25.161, df = 2, P < 0.01) revealed that there were significantly more eggs in Roanoke samples (41.29 2.65) than Milford (29.40 2.04; Dunn test P < 0.01) and Winchester (26.47 2.08; Dunn test P < 0.01).

```

## Kaplan-Meier

```{r}

#generate Kaplan-Meier curves
hatch_surv_obj <- with(hatch, Surv(time = right, time2 = left, type = "interval2"))
head(hatch_surv_obj, 100)
hatch_km_fit <- survfit(hatch_surv_obj ~ sitemonth + run, data = hatch)
hatch_km_fit ##can see LT50

#exract data
raw_hatch_survd <- data.frame(temp = hatch_km_fit$time,
                        propalive = hatch_km_fit$surv, 
                        lower = hatch_km_fit$lower, 
                        upper = hatch_km_fit$upper)
hatch_survd_newdata <- data.frame(strata = hatch_km_fit$strata)
hatch_survd_newdata <- hatch_survd_newdata %>%
  rownames_to_column(var = "var") %>%
  separate(col = "var", 
           into = c("sitemonth", "run"),
           sep = ",") %>%
  mutate(sitemonth = sub(".*=", "", sitemonth)) %>%
  mutate(run = sub(".*=", "", run)) %>%
  slice(rep(row_number(), strata))
hatch_survd <- cbind(raw_hatch_survd, hatch_survd_newdata[,-3]) %>%
  mutate(
    sitemonth = str_replace_all(sitemonth, "\\s+", ""),
    run = str_replace_all(run, "\\s+", ""),
  propalive = 1-propalive, #take the inverse, as to display survival not mortality
  lower = 1-lower, #take the inverse
  upper = 1-upper) #take the inverse

hatch_survdOG <- hatch_survd #create version before removing extra 1s and changing 1s to .999 and 0s to 0.001

#plot
ggplot()+
  geom_point(data = hatch_survd, aes(x = temp, y = propalive, color = run))+
  geom_line(data = hatch_survd, aes(x = temp, y = propalive, color = run))+
  geom_ribbon(data = hatch_survd, aes(x = temp, ymin = lower, ymax = upper, fill = run), alpha = 0.1)+
  facet_wrap(sitemonth~., nrow=2)+
  theme(legend.position = c(.85,.25))+
  scale_x_reverse()

```

## data prep

```{r}

abbott_hatch1sMF <- hatch_survd %>%
  group_by(sitemonth, run) %>%
  mutate(max = max(propalive)) %>%
    mutate(propalive_abbott = propalive/max,
           upper_abbott = lower/max,
           lower_abbott = upper/max) #apply Abbott's correction

abbott_hatch1s <- abbott_hatch1sMF %>%filter(!(sitemonth == "MilfordFebruary" & run == "first")) #would have to get rid of MilFeb1 since there was no hatch at all

abbott_hatch <- abbott_hatch1s[-c(10,11,12,20,21,22,29,30,31,37,38,39,47,54,55,56,62,63,64,72,73,74,75,76,86,94,103,111,117,121,129,145,146,153,154,155), ] #remove all 1s except for the 1 with the lowest x valye within each site*month*run

abbott_hatch <- abbott_hatch %>% 
  mutate(propalive_abbott = ifelse(propalive_abbott == 0, 0.001, ifelse(propalive_abbott >0.998, 0.999, propalive_abbott)), #have to do this again then because we've created 1s
         upper_abbott = ifelse(upper_abbott >1,1,upper_abbott))
max(abbott_hatch$propalive_abbott)
#plot
ggplot()+
  geom_point(data = abbott_hatch, aes(x = temp, y = propalive_abbott, color = run))+
  geom_line(data = abbott_hatch, aes(x = temp, y = propalive_abbott, color = run)) +
  geom_ribbon(data = abbott_hatch, aes(x = temp, ymin = lower_abbott, ymax = upper_abbott, fill = run), alpha = 0.1)+
  facet_wrap(sitemonth~., nrow=2)+
  scale_x_reverse()+
  theme(legend.position = c(.85,.25))

```

## beta reg with sitemonth

```{r}
abbott_hatch$sitemonth <- factor(abbott_hatch$sitemonth, levels = c("MilfordDecember","MilfordJanuary","MilfordFebruary","RoanokeDecember","RoanokeJanuary","WinchesterDecember","WinchesterJanuary" ))
hatch_LLTempModel <- glmmTMB(propalive_abbott ~ temp * sitemonth + (1|run), data = abbott_hatch, family = beta_family(link = "logit"))
summary(glht(hatch_LLTempModel))

abbott_hatch$sitemonth <- factor(abbott_hatch$sitemonth, levels = c("MilfordJanuary","MilfordFebruary","RoanokeDecember","RoanokeJanuary","WinchesterDecember","WinchesterJanuary","MilfordDecember" ))
hatch_LLTempModel <- glmmTMB(propalive_abbott ~ temp * sitemonth + (1|run), data = abbott_hatch, family = beta_family(link = "logit"))
summary(glht(hatch_LLTempModel))

abbott_hatch$sitemonth <- factor(abbott_hatch$sitemonth, levels = c("MilfordFebruary","MilfordJanuary","RoanokeDecember","RoanokeJanuary","WinchesterDecember","WinchesterJanuary","MilfordDecember" ))
hatch_LLTempModel <- glmmTMB(propalive_abbott ~ temp * sitemonth + (1|run), data = abbott_hatch, family = beta_family(link = "logit"))
summary((hatch_LLTempModel))

abbott_hatch$sitemonth <- factor(abbott_hatch$sitemonth, levels = c("WinchesterDecember","WinchesterJanuary","MilfordDecember","MilfordFebruary","MilfordJanuary","RoanokeDecember","RoanokeJanuary" ))
hatch_LLTempModel <- glmmTMB(propalive_abbott ~ temp * sitemonth + (1|run), data = abbott_hatch, family = beta_family(link = "logit"))
summary(glht(hatch_LLTempModel))

abbott_hatch$sitemonth <- factor(abbott_hatch$sitemonth, levels = c("WinchesterJanuary","MilfordDecember","MilfordFebruary","MilfordJanuary","RoanokeDecember","RoanokeJanuary","WinchesterDecember" ))
hatch_LLTempModel <- glmmTMB(propalive_abbott ~ temp * sitemonth + (1|run), data = abbott_hatch, family = beta_family(link = "logit"))
summary((hatch_LLTempModel))

abbott_hatch$sitemonth <- factor(abbott_hatch$sitemonth, levels = c("RoanokeDecember","WinchesterJanuary","MilfordDecember","MilfordFebruary","MilfordJanuary","RoanokeJanuary","WinchesterDecember" ))
hatch_LLTempModel <- glmmTMB(propalive_abbott ~ temp * sitemonth + (1|run), data = abbott_hatch, family = beta_family(link = "logit"))
summary(glht(hatch_LLTempModel))

abbott_hatch$sitemonth <- factor(abbott_hatch$sitemonth, levels = c("RoanokeJanuary","RoanokeDecember","WinchesterJanuary","MilfordDecember","MilfordFebruary","MilfordJanuary","WinchesterDecember" ))
hatch_LLTempModel <- glmmTMB(propalive_abbott ~ temp * sitemonth + (1|run), data = abbott_hatch, family = beta_family(link = "logit"))
summary(glht(hatch_LLTempModel))

#sample sizes
table(hatch[!(hatch$sitemonth == "MilfordFebruary" & hatch$run == "first"),]$sitemonth) #396-814
table(abbott_hatch$sitemonth)

df.residual(hatch_LLTempModel)
#calculate proportion variance explained by random effects
var <- var(residuals(hatch_LLTempModel))
randovar <- 8.672e-10

summary((hatch_LLTempModel))
randovar/(randovar+var)
#1.883615e-08

#find the dispersion parameter theta/phi
sigma(hatch_LLTempModel) #1.21

```

## plot and quantiles with month

```{r}
#make predictions
newdata <- expand.grid(sitemonth = c("MilfordDecember","MilfordJanuary","MilfordFebruary","WinchesterDecember","WinchesterJanuary", "RoanokeDecember","RoanokeJanuary"), temp = c(seq(10, -5, by = -1), seq(-5, -45, by = -0.01))) 
hatch_LLTempP <- cbind(newdata,
                 predict(hatch_LLTempModel, newdata, type = "response", se.fit = T, re.form = NA))
hatch_LLTempP$lower <- hatch_LLTempP$fit - 1.96 * hatch_LLTempP$se.fit
hatch_LLTempP$upper <- hatch_LLTempP$fit + 1.96 * hatch_LLTempP$se.fit

#retrieve quantiles
calculate_quantiles <- function(data) {
  list(
    data.frame(
      quantile = "50%", 
      estimate = data[which.min(abs(data$fit - 0.5)), ]$temp,
      lower = data[which.min(abs(data$lower - 0.5)), ]$temp, 
      upper = data[which.min(abs(data$upper - 0.5)), ]$temp
    ),
    data.frame(
      quantile = "90%", 
      estimate = data[which.min(abs(data$fit - 0.1)), ]$temp,
      lower = data[which.min(abs(data$lower - 0.1)), ]$temp, 
      upper = data[which.min(abs(data$upper - 0.1)), ]$temp
    )
  )
}
hatch_quantiles <- hatch_LLTempP %>%
  group_by(sitemonth) %>%
  do(bind_rows(calculate_quantiles(.))) %>%
  ungroup()
hatch_quantiles[hatch_quantiles$quantile == "50%",]

write.csv(hatch_quantiles, file = "C:/Users/CLAHM/Downloads/SLF/Data/SLF_LLTemp_hatch_quantiles.csv")

#plot
hatch_LLTempP <- hatch_LLTempP %>%
  mutate(lower2 = ifelse(lower<0,0,lower),
         upper2 = ifelse(upper>1,1,upper))
hatch_LLTempP$sitemonth <- factor(hatch_LLTempP$sitemonth, levels = c("MilfordDecember","MilfordJanuary","MilfordFebruary","WinchesterDecember","WinchesterJanuary","RoanokeDecember","RoanokeJanuary"))
abbott_hatch$sitemonth <- factor(abbott_hatch$sitemonth, levels = c("MilfordDecember","MilfordJanuary","MilfordFebruary","WinchesterDecember","WinchesterJanuary","RoanokeDecember","RoanokeJanuary"))

ggplot() +
  geom_point(data = abbott_hatch, aes(x = temp, y = propalive_abbott, color = run)) +
  geom_errorbar(data = abbott_hatch, aes(x = temp, ymin = lower_abbott, ymax = upper_abbott, color = run), alpha = 0.5) +
  geom_line(data = hatch_LLTempP, aes(x = temp, y = fit)) +
  geom_ribbon(data = hatch_LLTempP, aes(x = temp, ymin = lower2, ymax = upper2), alpha = 0.1)+
  facet_wrap2(sitemonth~., axes = "all", remove_labels = "all", nrow=2)+
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line(),
    legend.position = c(.85, .25)) +
  labs(color = "", fill = "", x = "Temperature (°C)", y = "Proportion Alive")+
  scale_x_reverse(limits = c(5, -60))

#overlayed plot
ggplot() +
  geom_point(data = abbott_hatch, aes(x = temp, y = propalive_abbott, color = sitemonth)) +
  geom_errorbar(data = abbott_hatch, aes(x = temp, ymin = lower_abbott, ymax = upper_abbott, color = sitemonth), alpha=0.5) +
  geom_line(data = hatch_LLTempP, aes(x = temp, y = fit, color = sitemonth)) +
  geom_ribbon(data = hatch_LLTempP, aes(x = temp, ymin = lower2, ymax = upper2, fill = sitemonth), alpha = 0.1)+
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line()) +
  labs(color = "", fill = "", x = "Temperature (°C)", y = "Proportion Alive")+
  scale_x_reverse()+
  scale_fill_manual(values = Palette)+
  scale_color_manual(values = Palette)

```

## beta reg without month

```{r}
#make model
abbott_hatch$run2 <- paste(abbott_hatch$sitemonth, abbott_hatch$run, sep = "")

hatch_LLTempModel_pool <- glmmTMB(propalive_abbott ~ temp + (1|run2), data = abbott_hatch, family = beta_family(link = "logit"))
summary((hatch_LLTempModel_pool))

df.residual(hatch_LLTempModel_pool)
#calculate proportion variance explained by random effects
var <- var(residuals(hatch_LLTempModel_pool))
var
randovar <- 3.524e-09
randovar/(randovar+var)
#1.883615e-08

#find the dispersion parameter theta/phi
sigma(hatch_LLTempModel_pool) #0.969
```

## plot and quantiles without month

```{r}
#make predictions
newdata <- data.frame(temp = c(seq(10, -5, by = -1), seq(-5, -45, by = -0.01))) 
hatch_LLTempP_pool <- cbind(newdata,
                 predict(hatch_LLTempModel_pool, newdata, type = "response", se.fit = T, re.form = NA))
hatch_LLTempP_pool$lower <- hatch_LLTempP_pool$fit - 1.96 * hatch_LLTempP_pool$se.fit
hatch_LLTempP_pool$upper <- hatch_LLTempP_pool$fit + 1.96 * hatch_LLTempP_pool$se.fit

#retrieve quantiles
quantiles_pool <- rbind(data.frame(quantile = "50%", 
      estimate = hatch_LLTempP_pool[which.min(abs(hatch_LLTempP_pool$fit - 0.5)), ]$temp,
      lower = hatch_LLTempP_pool[which.min(abs(hatch_LLTempP_pool$lower - 0.5)), ]$temp, 
      upper = hatch_LLTempP_pool[which.min(abs(hatch_LLTempP_pool$upper - 0.5)), ]$temp),
    data.frame(quantile = "90%", 
      estimate = hatch_LLTempP_pool[which.min(abs(hatch_LLTempP_pool$fit - 0.1)), ]$temp,
      lower = hatch_LLTempP_pool[which.min(abs(hatch_LLTempP_pool$lower - 0.1)), ]$temp, 
      upper = hatch_LLTempP_pool[which.min(abs(hatch_LLTempP_pool$upper - 0.1)), ]$temp))
quantiles_pool


hatch_LLTempP_pool <- hatch_LLTempP_pool %>%
  mutate(lower2 = ifelse(lower<0,0,lower),
         upper2 = ifelse(upper>1,1,upper))

ggplot() +
  geom_point(data = abbott_hatch, aes(x = temp, y = propalive_abbott)) +
  geom_line(data = hatch_LLTempP_pool, aes(x = temp, y = fit)) +
  geom_ribbon(data = hatch_LLTempP_pool, aes(x = temp, ymin = lower2, ymax = upper2), alpha = 0.1)+
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line()) +
  labs(color = "", fill = "", x = "Temperature(°C)", y = "Proportion Alive")+
  scale_x_reverse()

```    

# BVT vs. HATCH

## combo beta reg

```{r}
combo_bvt <- survd[, c("temp", "propalive", "lower", "upper", "sitemonth", "run2"), ]
combo_hatch <- abbott_hatch[, c("temp", "propalive_abbott", "lower_abbott", "upper_abbott", "sitemonth", "run2")]
names(combo_hatch) <- c("temp", "propalive", "lower", "upper", "sitemonth", "run2")
combo_bvt$Method <- "BVT"
combo_hatch$Method <- "hatch"

combo_methods <- rbind(combo_bvt, combo_hatch)

combo_methods$type <- factor(combo_methods$Method, levels = c("hatch", "BVT"))

combo_methods_model <- glmmTMB(propalive ~ temp * Method + (1|run2), data = combo_methods, family = beta_family(link = "logit"))
summary(glht(combo_methods_model))

#sample sizes
table(combo_methods$Method)
df.residual(combo_methods_model)

#calculate proportion variance explained by random effects
var <- var(residuals(combo_methods_model))
randovar <- 8.909e-10
var
randovar/(randovar+var)

#find the dispersion parameter theta/phi
sigma(combo_methods_model)

```

## combo plot and quantiles

```{r}

#make predictions
newdata <- expand.grid(Method = c("BVT", "hatch"), temp = c(seq(10, -5, by = -1), seq(-5, -45, by = -0.01)))

combo_methodsP <- cbind(newdata,
                 predict(combo_methods_model, newdata, type = "response", se.fit = T, re.form = NA))

combo_methodsP$lower <- combo_methodsP$fit - 1.96 * combo_methodsP$se.fit
combo_methodsP$upper <- combo_methodsP$fit + 1.96 * combo_methodsP$se.fit

#retrieve quantiles
calculate_quantiles <- function(data) {
  list(
    data.frame(
      quantile = "50%", 
      estimate = data[which.min(abs(data$fit - 0.5)), ]$temp,
      lower = data[which.min(abs(data$lower - 0.5)), ]$temp, 
      upper = data[which.min(abs(data$upper - 0.5)), ]$temp
    ),
    data.frame(
      quantile = "90%", 
      estimate = data[which.min(abs(data$fit - 0.1)), ]$temp,
      lower = data[which.min(abs(data$lower - 0.1)), ]$temp, 
      upper = data[which.min(abs(data$upper - 0.1)), ]$temp
    )
  )
}
combo_methods_quantiles <- combo_methodsP %>%
  group_by(Method) %>%
  do(bind_rows(calculate_quantiles(.))) %>%
  ungroup()

#plot
combo_methodsP <- combo_methodsP %>%
  mutate(lower2 = ifelse(lower<0,0,lower),
         upper2 = ifelse(upper>1,1,upper))

ggplot() +
  geom_point(data = combo_methods, aes(x = temp, y = propalive, color = Method, shape = Method), alpha = 0.5) +
  geom_line(data = combo_methodsP, aes(x = temp, y = fit, color = Method, linetype = Method)) +
  geom_ribbon(data = combo_methodsP, aes(x = temp, ymin = lower2, ymax = upper2, fill = Method), alpha = 0.1)+
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line(),
    legend.position = c(.88,.9),
    plot.background = element_rect(fill = "white", color = NA),
    legend.background = element_rect(fill = "white", color = NA)) +
  labs(color = "Method", fill = "Method", linetype = "Method", shape = "Method", x = "Temperature (°C)", y = "Proportion Alive")+
  scale_x_reverse(limits = c(0, -45),breaks = c(0, -5, -10, -15, -20, -25, -30, -35, -40, -45))+
  scale_color_manual(values = c( "#0072B2","red"),
                     labels = c("BVT", "hatch")) +
  scale_fill_manual(values = c("#0072B2", "red"),
                     labels = c("BVT", "hatch"))+
  scale_shape_manual(values = c(19, 2),
                     labels = c("BVT", "hatch")) +
  scale_linetype_manual(values = c("solid", "dashed"), 
                        labels = c("BVT", "hatch"))


```

# SCP

## remove outliers

```{r}
#remove outliers
unique(SCP$sitemonth)
list_quantiles <- tapply(SCP$SCP, SCP$sitemonth, quantile)
Q1s <- sapply(1:9, function(i) list_quantiles[[i]][2])
Q3s <- sapply(1:9, function(i) list_quantiles[[i]][4])
IQRs <- tapply(SCP$SCP, SCP$sitemonth, IQR)

Lowers <- Q1s - 1.5*IQRs
Uppers <- Q3s + 1.5*IQRs
 Uppers
 Lowers
datas <- split(SCP, SCP$sitemonth)
Uppers[2]<- -23.89875 #February Milford is unusually high upper bound, so change it to the next highest; visually it makes sense

#SCPNO = no outliers
SCPNO <- NULL
for (i in 1:9){
out <- subset(datas[[i]], datas[[i]]$SCP < Uppers[i]) #don't want to remove the low outliers in SCPOs, so only removing upper outliers
SCPNO <- rbind(SCPNO, out)
}
#SCPOs = outliers that were removed
SCPOs <- NULL
for (i in 1:9){
outs <- subset(datas[[i]],datas[[i]]$SCP > Uppers[i])
SCPOs <- rbind(SCPOs, outs)
}

dim(SCP)
dim(SCPNO)
dim(SCPOs)

#range of removed outliers
min(SCPOs$SCP)
max(SCPOs$SCP)

min(SCP$SCP)

```

## lmer-preliminary

```{r}
shapiro.test(SCPNO[SCPNO$sitemonth == "WinchesterNovember", ]$SCP)
hist(SCPNO[SCPNO$sitemonth == "WinchesterNovember", ]$SCP)
shapiro.test(SCPNO[SCPNO$sitemonth == "RoanokeNovember", ]$SCP)
hist(SCPNO[SCPNO$sitemonth == "RoanokeNovember", ]$SCP)
shapiro.test(SCPNO[SCPNO$sitemonth == "MilfordDecember", ]$SCP)
hist(SCPNO[SCPNO$sitemonth == "MilfordDecember", ]$SCP)
shapiro.test(SCPNO[SCPNO$sitemonth == "WinchesterDecember", ]$SCP)
hist(SCPNO[SCPNO$sitemonth == "WinchesterDecember", ]$SCP)
shapiro.test(SCPNO[SCPNO$sitemonth == "RoanokeDecember", ]$SCP)
hist(SCPNO[SCPNO$sitemonth == "RoanokeDecember", ]$SCP)
shapiro.test(SCPNO[SCPNO$sitemonth == "MilfordJanuary", ]$SCP)
hist(SCPNO[SCPNO$sitemonth == "MilfordJanuary", ]$SCP)
shapiro.test(SCPNO[SCPNO$sitemonth == "WinchesterJanuary", ]$SCP)
hist(SCPNO[SCPNO$sitemonth == "WinchesterJanuary", ]$SCP)
shapiro.test(SCPNO[SCPNO$sitemonth == "RoanokeJanuary", ]$SCP)
hist(SCPNO[SCPNO$sitemonth == "RoanokeJanuary", ]$SCP)
shapiro.test(SCPNO[SCPNO$sitemonth == "MilfordFebruary", ]$SCP) #only not normal one
hist(SCPNO[SCPNO$sitemonth == "MilfordFebruary", ]$SCP)
unique(SCPNO$sitemonth)
SCPNO$mass2 <- paste(SCPNO$sitemonth, SCPNO$mass, sep = "")

#SCPNO$site <- factor(SCPNO$site, levels = c("Roanoke", "Winchester", "Milford", "Westport"))
SCPNOmodel <- lmer(SCP ~ sitemonth + (1|mass2), data = SCPNO)
summary(SCPNOmodel)

#SCPNO$sitemonth <- factor(SCPNO$sitemonth, levels = c("RoanokeNovember", "WinchesterNovember", "RoanokeDecember", "WinchesterDecember", "MilfordDecember", "RoanokeJanuary", "WinchesterJanuary", "MilfordJanuary", "MilfordFebruary"))
SCPNOmodel <- lmer(SCP ~ sitemonth + (1|mass), data = SCPNO)
summary(SCPNOmodel)

#find proportion of variance explained by random effect: egg mass
#SCPNONO$site <- factor(SCPNONO$site, levels = c("Winchester", "Roanoke", "Milford"))
0.01687/(0.01687+0.80847)

#SCPNO$sitemonth <- factor(SCPNO$sitemonth, levels = c("RoanokeNovember", "WinchesterNovember", "RoanokeDecember", "WinchesterDecember", "MilfordDecember", "RoanokeJanuary", "WinchesterJanuary", "MilfordJanuary", "MilfordFebruary"))
ggplot(SCPNO, aes(x = sitemonth, y = SCP)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(width = 0.1, alpha = 0.2) +
  labs(x = "", y = "Supercooling Point (°C)")+
  theme_minimal()+
  theme(panel.border = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line(),
    axis.text.x = element_text(angle = 30, hjust = 1))

```

## power analysis

```{r}
SCPmodel_simple <- lm(SCP ~ month, data = SCP[SCP$month %in% c("December", "January"),]) #interchange to whichever comparison interested in
summary(SCPmodel_simple)

r_squared <- summary(SCPmodel_simple)$r.squared
effect_size <- r_squared / (1 - r_squared)  # Cohen's f² (effect size)

pwr.f2.test(u = (length(coef(SCPmodel_simple)) - 1), f2 = effect_size, sig.level = 0.05, power = 0.80)

```

## Kaplan-Meier

```{r}
#generate Kaplan-Meier curves
SCPsurv_obj <- with(SCPNO, Surv(time = SCP, time2 = SCP, type = "interval2"))
head(SCPsurv_obj, 100)
SCPkm_fit <- survfit(SCPsurv_obj ~ sitemonth+mass, data = SCPNO)
SCPkm_fit
#exract data
SCPsurvd <- data.frame(temp = SCPkm_fit$time,
                        propSCP = SCPkm_fit$surv, 
                        lower = SCPkm_fit$lower, 
                        upper = SCPkm_fit$upper)
SCPsurvd_newdata <- data.frame(strata = SCPkm_fit$strata)

SCPsurvd_newdata <- SCPsurvd_newdata %>%
  rownames_to_column(var = "sitemonth") %>%
  separate(sitemonth, into = c("sitemonth", "mass"), sep = ", ")%>%
  mutate(sitemonth = sub(".*=", "", sitemonth),
         mass = sub(".*=", "", mass)) %>%
  slice(rep(row_number(), strata))
SCPsurvd <- cbind(SCPsurvd, SCPsurvd_newdata[,-3]) %>%
  mutate(propSCP = 1-propSCP, #take the inverse, as to display survival not mortality
  lower = 1-lower, #take the inverse
  upper = 1-upper) #take the inverse

ggplot()+
  geom_step(data = SCPsurvd, aes(x = temp, y = propSCP, color = mass))+
  geom_ribbon(data = SCPsurvd, aes(x = temp, ymin = lower, ymax = upper, fill = sitemonth), alpha = 0.1)+
  facet_wrap(sitemonth~.)+
  scale_x_reverse() 

SCPsurvd$propSCP <- ifelse(SCPsurvd$propSCP < 0.00001, 0.001, ifelse(SCPsurvd$propSCP == 1, 0.999, SCPsurvd$propSCP))

```

## beta reg with month

```{r}

SCPsurvd$sitemonth <- factor(SCPsurvd$sitemonth, levels = c("MilfordDecember", "WinchesterDecember", "RoanokeDecember", "RoanokeNovember", "WinchesterNovember","RoanokeJanuary", "WinchesterJanuary", "MilfordJanuary", "MilfordFebruary"))
SCPModel <- glmmTMB(propSCP~ temp* sitemonth + (1|mass), data = SCPsurvd, family = beta_family(link = "logit"))
summary(SCPModel)

#other references
SCPsurvd$sitemonth <- factor(SCPsurvd$sitemonth, levels = c("MilfordJanuary", "WinchesterJanuary", "RoanokeJanuary", "MilfordDecember", "WinchesterDecember","RoanokeDecember", "RoanokeNovember", "WinchesterNovember", "MilfordFebruary"))
SCPModel <- glmmTMB(propSCP~ temp* sitemonth + (1|mass), data = SCPsurvd, family = beta_family(link = "logit"))
summary(glht(SCPModel), adjust = "bonferroni")

SCPsurvd$sitemonth <- factor(SCPsurvd$sitemonth, levels = c("MilfordFebruary", "MilfordJanuary", "WinchesterJanuary", "RoanokeJanuary", "MilfordDecember","WinchesterDecember", "RoanokeDecember", "RoanokeNovember", "WinchesterNovember"))
SCPModel <- glmmTMB(propSCP~ temp* sitemonth + (1|mass), data = SCPsurvd, family = beta_family(link = "logit"))
summary(glht(SCPModel), adjust = "bonferroni")

SCPsurvd$sitemonth <- factor(SCPsurvd$sitemonth, levels = c("WinchesterNovember", "RoanokeDecember", "WinchesterDecember", "MilfordDecember","RoanokeJanuary", "WinchesterJanuary", "MilfordJanuary", "MilfordFebruary", "RoanokeNovember"))
SCPModel <- glmmTMB(propSCP~ temp* sitemonth + (1|mass), data = SCPsurvd, family = beta_family(link = "logit"))
summary(SCPModel)

SCPsurvd$sitemonth <- factor(SCPsurvd$sitemonth, levels = c("WinchesterDecember", "RoanokeDecember", "RoanokeNovember", "WinchesterNovember", "MilfordDecember","RoanokeJanuary", "WinchesterJanuary", "MilfordJanuary", "MilfordFebruary"))
SCPModel <- glmmTMB(propSCP~ temp* sitemonth + (1|mass), data = SCPsurvd, family = beta_family(link = "logit"))
summary(SCPModel)

SCPsurvd$sitemonth <- factor(SCPsurvd$sitemonth, levels = c("WinchesterJanuary", "RoanokeJanuary", "MilfordDecember", "WinchesterDecember", "RoanokeDecember","RoanokeNovember", "WinchesterNovember", "MilfordJanuary", "MilfordFebruary"))
SCPModel <- glmmTMB(propSCP~ temp* sitemonth + (1|mass), data = SCPsurvd, family = beta_family(link = "logit"))
summary(SCPModel)

SCPsurvd$sitemonth <- factor(SCPsurvd$sitemonth, levels = c("RoanokeNovember", "WinchesterNovember", "RoanokeDecember", "WinchesterDecember", "MilfordDecember","RoanokeJanuary", "WinchesterJanuary", "MilfordJanuary", "MilfordFebruary"))
SCPModel <- glmmTMB(propSCP~ temp* sitemonth + (1|mass), data = SCPsurvd, family = beta_family(link = "logit"))
summary((SCPModel))

SCPsurvd$sitemonth <- factor(SCPsurvd$sitemonth, levels = c("RoanokeDecember", "RoanokeNovember", "WinchesterNovember", "WinchesterDecember", "MilfordDecember","RoanokeJanuary", "WinchesterJanuary", "MilfordJanuary", "MilfordFebruary"))
SCPModel <- glmmTMB(propSCP~ temp* sitemonth + (1|mass), data = SCPsurvd, family = beta_family(link = "logit"))
summary(SCPModel)

SCPsurvd$sitemonth <- factor(SCPsurvd$sitemonth, levels = c("RoanokeJanuary", "MilfordDecember", "WinchesterDecember", "RoanokeDecember", "RoanokeNovember","WinchesterNovember", "WinchesterJanuary", "MilfordJanuary", "MilfordFebruary"))
SCPModel <- glmmTMB(propSCP~ temp* sitemonth + (1|mass), data = SCPsurvd, family = beta_family(link = "logit"))
summary(SCPModel)

#sample sizes
SCPsurvd$sitemonth <- factor(SCPsurvd$sitemonth, levels = c("MilfordDecember", "MilfordJanuary", "MilfordFebruary", "WinchesterNovember", "WinchesterDecember", "WinchesterJanuary", "RoanokeNovember","RoanokeDecember","RoanokeJanuary"))
SCPNO$sitemonth <- factor(SCPNO$sitemonth, levels = c("MilfordDecember", "MilfordJanuary", "MilfordFebruary", "WinchesterNovember", "WinchesterDecember", "WinchesterJanuary", "RoanokeNovember","RoanokeDecember","RoanokeJanuary"))
table(SCPNO$sitemonth)
table(SCPsurvd$sitemonth)

#calculate proportion variance explained by random effects
var <- var(residuals(SCPModel))
randovar <- 0.1125

var
randovar/(randovar+var)


#find the dispersion parameter theta/phi
sigma(SCPModel)

```

## plot and quantiles with month

```{r}
#make predictions
SCPsurvd$sitemonth <- factor(SCPsurvd$sitemonth, levels = c("MilfordDecember", "MilfordJanuary", "MilfordFebruary", "WinchesterNovember", "WinchesterDecember","WinchesterJanuary", "RoanokeNovember", "RoanokeDecember", "RoanokeJanuary"))
SCPModel <- glmmTMB(propSCP~ temp* sitemonth + (1|mass), data = SCPsurvd, family = beta_family(link = "logit"))
summary(SCPModel)

newdata <- expand.grid(sitemonth = c("MilfordDecember", "MilfordJanuary", "MilfordFebruary", "WinchesterNovember", "WinchesterDecember","WinchesterJanuary", "RoanokeNovember", "RoanokeDecember", "RoanokeJanuary"), temp = c(seq(10, -12, by = -1), seq(-12, -32, by = -0.01), seq(-32, -55, by = -1)))

scpP <- cbind(newdata,
                 predict(SCPModel, newdata, type = "response", se.fit = T, re.form = NA))
# scpP <- scpP %>%
#   mutate(month = str_split(sitemonth, " ", simplify = TRUE)[, 1],
#          site = str_split(sitemonth, " ", simplify = TRUE)[, 2])
scpP$lower <- scpP$fit - 1.96 * scpP$se.fit
scpP$upper <- scpP$fit + 1.96 * scpP$se.fit

#retrieve quantiles
calculate_quantiles <- function(data) {
  list(
    data.frame(
      quantile = "50%", 
      estimate = data[which.min(abs(data$fit - 0.5)), ]$temp,
      lower = data[which.min(abs(data$lower - 0.5)), ]$temp, 
      upper = data[which.min(abs(data$upper - 0.5)), ]$temp
    ),
    data.frame(
      quantile = "90%", 
      estimate = data[which.min(abs(data$fit - 0.9)), ]$temp,
      lower = data[which.min(abs(data$lower - 0.9)), ]$temp, 
      upper = data[which.min(abs(data$upper - 0.9)), ]$temp
    )
  )
}
scp_quantiles <- scpP %>%
  group_by(sitemonth) %>%
  do(bind_rows(calculate_quantiles(.))) %>%
  ungroup()
scp_quantiles[scp_quantiles$quantile == "50%",]
write.csv(scp_quantiles, file = "C:/Users/CLAHM/Downloads/SLF/Data/SLF_SCP_quantiles.csv")

#plot
scpP <- scpP %>%
  mutate(lower2 = ifelse(lower<0,0,lower),
         upper2 = ifelse(upper>1,1,upper))
SCPsurvd$sitemonth <- factor(SCPsurvd$sitemonth, levels = c("MilfordDecember", "MilfordJanuary", "MilfordFebruary", "WinchesterNovember", "WinchesterDecember", "WinchesterJanuary", "RoanokeNovember","RoanokeDecember","RoanokeJanuary"))
scpP$sitemonth <- factor(scpP$sitemonth, levels = c("MilfordDecember", "MilfordJanuary", "MilfordFebruary", "WinchesterNovember", "WinchesterDecember", "WinchesterJanuary", "RoanokeNovember","RoanokeDecember","RoanokeJanuary"))

ggplot() +
  geom_point(data = SCP, aes(x = temp, y = propSCP, color = sitemonth)) +
  geom_line(data = scpP, aes(x = temp, y = fit, color = sitemonth)) +
  geom_ribbon(data = scpP, aes(x = temp, ymin = lower2, ymax = upper2, fill = sitemonth), alpha = 0.1)+
  scale_x_reverse(limits = c(-20, -32)) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line(),
    legend.position.inside()) +
  labs(color = "", fill = "", x = "Temperature(°C)", y = "Proportion Supercooled")+
  scale_color_manual(values = Palette)+
  scale_fill_manual(values = Palette)+
  scale_y_reverse(breaks = c(1, 0.9, 0.75, 0.5, 0.25, 0), labels = c("0", "0.1","0.25","0.5","0.75","1"))

#Clarissa's original of above code:
#ggplot() +
#  geom_point(data = SCPsurvd, aes(x = temp, y = propSCP, color = sitemonth)) +
#  geom_line(data = scpP, aes(x = temp, y = fit, color = sitemonth)) +
#  geom_ribbon(data = scpP, aes(x = temp, ymin = lower2, ymax = upper2, fill = sitemonth), alpha = 0.1)+
#  scale_x_reverse(limits = c(-20, -32)) +
#  theme_minimal() +
#   theme(panel.grid.minor = element_blank(),
#    axis.line = element_line(),
#    axis.ticks = element_line()) +
#  labs(color = "", fill = "", x = "Temperature(°C)", y = "Proportion Supercooled")+
#  scale_color_manual(values = Palette)+
#  scale_fill_manual(values = Palette)+
#  scale_y_reverse(breaks = c(1, 0.9, 0.75, 0.5, 0.25, 0), labels = c("0", "0.1","0.25","0.5","0.75","1"))

```

## beta reg without month

```{r}
#make model
SCPsurvd$mass2 <- paste(SCPsurvd$sitemonth, SCPsurvd$mass, sep = "")
scpModel_pool <- glmmTMB(propSCP ~ temp + (1|mass2), data = SCPsurvd, family = beta_family(link = "logit"))
summary(scpModel_pool)

#calculate proportion variance explained by random effects
var <- var(residuals(scpModel_pool))
randovar <- 0.4397
var
randovar/(randovar+var)
#1.332563e-08 or 0.00000001332563 or 0.00000133

#find the dispersion parameter theta/phi
sigma(scpModel_pool)

```

## plot and quantiles without month

```{r}
#make predictions
newdata <- data.frame(temp = c(seq(10, -12, by = -1), seq(-12, -40, by = -0.01), seq(-40, -55, by = -1))) 
scpP_pool <- cbind(newdata,
                 predict(scpModel_pool, newdata, type = "response", se.fit = T, re.form = NA))
scpP_pool$lower <- scpP_pool$fit - 1.96 * scpP_pool$se.fit
scpP_pool$upper <- scpP_pool$fit + 1.96 * scpP_pool$se.fit

scpP_pool <- scpP_pool %>%
  mutate(lower2 = ifelse(lower<0,0,lower),
         upper2 = ifelse(upper>1,1,upper))

#retrieve quantiles
quantiles_pool <- rbind(data.frame(quantile = "50%", 
      estimate = scpP_pool[which.min(abs(scpP_pool$fit - 0.5)), ]$temp,
      lower = scpP_pool[which.min(abs(scpP_pool$lower - 0.5)), ]$temp, 
      upper = scpP_pool[which.min(abs(scpP_pool$upper - 0.5)), ]$temp),
    data.frame(quantile = "90%", 
      estimate = scpP_pool[which.min(abs(scpP_pool$fit - 0.9)), ]$temp,
      lower = scpP_pool[which.min(abs(scpP_pool$lower - 0.9)), ]$temp, 
      upper = scpP_pool[which.min(abs(scpP_pool$upper - 0.9)), ]$temp))
quantiles_pool

#plot
ggplot() +
  geom_point(data = SCPsurvd, aes(x = temp, y = propSCP)) +
  geom_line(data = scpP_pool, aes(x = temp, y = fit)) +
  geom_ribbon(data = scpP_pool, aes(x = temp, ymin = lower2, ymax = upper2), alpha = 0.1)+
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line()) +
  labs(color = "", fill = "", x = "Temperature(°C)", y = "Proportion Supercooled")+
  scale_x_reverse()+
  scale_y_reverse(breaks = c(1, 0.9, 0.75, 0.5, 0.25, 0), labels = c("0", "0.1","0.25","0.5","0.75","1"))

```    

# LLTEMP VS SCP

## combo beta reg

```{r}
#combine dataframes
colnames(survd)
colnames(SCPsurvd)
survd_clean<-survd
SCPsurvd_clean <- SCPsurvd
names(survd_clean) <- c("temp", "prop", "lower", "upper", "sitemonth","mass")
names(SCPsurvd_clean) <- c("temp", "prop", "lower", "upper", "sitemonth","mass")
survd_clean$datatype <- "LLTemp"
SCPsurvd_clean$datatype <- "SCP"
colnames(survd_clean)
colnames(SCPsurvd_clean)
combo <- rbind(survd_clean, SCPsurvd_clean)

#model
combo$datatype <- factor(combo$datatype, levels = c("LLTemp", "SCP"))
combomodel <- glmmTMB(prop ~ temp * datatype + (1|mass), data = combo, family = beta_family(link = "logit"))
summary(combomodel)

combo$datatype <- factor(combo$datatype, levels = c("SCP", "LLTemp"))
combomodel <- glmmTMB(prop ~ temp * datatype + (1|mass), data = combo, family = beta_family(link = "logit"))
summary(combomodel)

#sample sizes
table(combo$datatype)
table(SCPsurvd$sitemonth)

#calculate proportion variance explained by random effects
var <- var(residuals(combomodel))
randovar <- 0.2043
var
randovar/(randovar+var)

#find the dispersion parameter theta/phi
sigma(combomodel)
```

## combo plot and quantiles

```{r}

#make predictions
newdata <- expand.grid(datatype = c("LLTemp", "SCP"), temp = c(seq(10, -12, by = -1), seq(-12, -42, by = -0.01), seq(-42, -55, by = -1)))
comboP <- cbind(newdata,
                 predict(combomodel, newdata, type = "response", se.fit = T, re.form = NA))
comboP$lower <- comboP$fit - 1.96 * comboP$se.fit
comboP$upper <- comboP$fit + 1.96 * comboP$se.fit

#retrieve quantiles
calculate_quantiles <- function(data) {
  list(
    data.frame(
      quantile = "50%", 
      estimate = data[which.min(abs(data$fit - 0.5)), ]$temp,
      lower = data[which.min(abs(data$lower - 0.5)), ]$temp, 
      upper = data[which.min(abs(data$upper - 0.5)), ]$temp
    ),
    data.frame(
      quantile = "90%", 
      estimate = data[which.min(abs(data$fit - 0.9)), ]$temp,
      lower = data[which.min(abs(data$lower - 0.9)), ]$temp, 
      upper = data[which.min(abs(data$upper - 0.9)), ]$temp
    )
  )
}
combo_quantiles <- comboP %>%
  group_by(type) %>%
  do(bind_rows(calculate_quantiles(.))) %>%
  ungroup()

#plot
comboP <- comboP %>%
  mutate(lower2 = ifelse(lower<0,0,lower),
         upper2 = ifelse(upper>1,1,upper))

ggplot() +
  geom_point(data = combo, aes(x = temp, y = prop, color = datatype, shape = datatype), alpha = 0.5) +
  geom_line(data = comboP, aes(x = temp, y = fit, color = datatype, linetype = datatype)) +
  geom_ribbon(data = comboP, aes(x = temp, ymin = lower2, ymax = upper2, fill = datatype), alpha = 0.1)+
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line(),
    legend.position = c(.89,.9),
    legend.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = NULL)) +
  labs(color = "", fill = "", linetype = "", shape = "", x = "Temperature (°C)", y = "Proportion Alive/Pre-Supercool")+
  scale_x_reverse(limits = c(0, -45),breaks = c(0, -5, -10, -15, -20, -25, -30, -35, -40, -45))+
  scale_color_manual(values = c( "#0072B2","red"),
                     labels = c("LLTemp", "SCP")) +
  scale_fill_manual(values = c("#0072B2", "red"),
                     labels = c("LLTemp", "SCP"))+
  scale_shape_manual(values = c(19, 2),
                     labels = c("LLTemp", "SCP")) +
  scale_linetype_manual(values = c("solid", "dashed"), 
                        labels = c("LLTemp", "SCP"))

```

## SCP/BVT/hatch plot

```{r}
min(LLTemp$temp)
min(SCP$SCP)#find minimum data value so i know where to start dashed lines indicating extrapolation

#combine datasets
# colnames(comboP)
# colnames(combo_methodsP)
# comboP$Method <- ifelse(comboP$datatype == "LLTemp", "BVT", "SCP")
# SCP_BVT_hatchP <- rbind(comboP[, c("temp", "fit", "lower2", "upper2", "Method")], combo_methodsP[, c("temp", "fit", "lower2", "upper2", "Method")])
# SCP_BVT_hatchP$lowest <- ifelse(SCP_BVT_hatchP$Method == "SCP", -29.64, -33.71)
colnames(LLTempP_pool)
colnames(hatch_LLTempP_pool)
colnames(scpP_pool)
LLTempP_pool$Method <- "BVT"
hatch_LLTempP_pool$Method <- "hatch"
scpP_pool$Method <- "SCP"
SCP_BVT_hatchP <- rbind(LLTempP_pool, hatch_LLTempP_pool, scpP_pool)
SCP_BVT_hatchP$lowest <- ifelse(SCP_BVT_hatchP$Method == "SCP", -29.64, -33.71)
                        
colnames(combo)
colnames(combo_methods)
combo$Method <- ifelse(combo$datatype == "LLTemp", "BVT", "SCP")
combo_methods <- combo_methods %>%
  mutate(prop2 = propalive)
colnames(combo)
colnames(combo_methods)
SCP_BVT_hatch <- rbind(combo[, c("temp", "prop", "lower", "upper", "sitemonth", "Method")], combo_methods[, c("temp", "prop", "lower", "upper", "sitemonth", "Method")])

#order methods
SCP_BVT_hatch$Method<- factor(SCP_BVT_hatch$Method, levels = c("hatch", "BVT", "SCP"))
SCP_BVT_hatchP$Method<- factor(SCP_BVT_hatchP$Method, levels = c("hatch", "BVT", "SCP"))

SCP_BVT_hatch_plot<- ggplot() +
  geom_point(data = SCP_BVT_hatch, aes(x = temp, y = prop, color = Method, shape = Method), alpha = 0.4, size = 0.8)+
  geom_line(data = SCP_BVT_hatchP, aes(x = temp, y = fit, color = Method, linetype = factor(temp > lowest))) +
  geom_ribbon(data = SCP_BVT_hatchP, aes(x = temp, ymin = lower2, ymax = upper2, fill = Method), alpha = 0.1)+
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line(),
    legend.position = c(.89,.82),
    legend.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white"),
    plot.margin = margin(10, 10, 5, 5)) +
  labs(x = "Temperature (°C)", y = "Proportion alive/pre-SCP")+
  scale_x_reverse(expand = c(0, 0), limits = c(0, -45),breaks = c(0, -5, -10, -15, -20, -25, -30, -35, -40, -45))+
  scale_y_continuous(expand = c(0.01, 0))+
  scale_color_manual(values = c( "#0072B2","#E69F00","red")) +
  scale_fill_manual(values = c( "#0072B2","#E69F00", "red"))+
  scale_shape_manual(values = c(2, 15, 19)) +
  scale_linetype_manual(values = c("dashed", "solid"))+
  guides(linetype = "none")

#errorbar line, to re-include it:
# +
#   geom_errorbar(data = SCP_BVT_hatch, aes(x = temp, ymin = lower, ymax = upper, color = Method), alpha = 0.3)

#save plot
  ggsave("C:/Users/venet001/Documents/Entomology/Venette/SLF/Outputs_from_R_Clarissa/Plots/SCP_BVT_hatch_plot.tiff", SCP_BVT_hatch_plot, width = 80, height = 80, units = "mm", dpi = 300)
#Clarissa's original below (line above adjusted to correct directory)
#ggsave("C:/Users/CLAHM/Downloads/SLF/Plots/SCP_BVT_hatch_plot.tiff", SCP_BVT_hatch_plot, width = 80, height = 80, units = "mm", dpi = 300)

```

