---
title: "SLF_2026 summaries"
author: "Clarissa Moore"
date: ""
output: html_document
---
 
## load data

```{r}
#change the file directories to yours
# SCP <- read.csv("D:/SLF/Data/SLF_SCP_data.csv")
# LLTemp <- read.csv("D:/SLF/Data/SLF_LLTemp_BVT_data.csv")
# hatch <- read.csv("D:/SLF/Data/SLF_LLTemp_hatch_data.csv")
#ok

#Clarissa's original statements were
SCP <- read.csv("C:/Users/CLAHM/Downloads/SLF/Data/SLF_SCP_data.csv")
LLTemp <- read.csv("C:/Users/CLAHM/Downloads/SLF/Data/SLF_LLTemp_BVT_data.csv")
LLTime <- read.csv("C:/Users/CLAHM/Downloads/SLF/Data/SLF_LLTime_data.csv")

```

## load packages

```{r, warning=FALSE, message=FALSE, results='hide'}
#this will download any packages if you don't already have them
packages <- c("tidyverse", "betareg", "survival", "lmerTest", "emmeans", "multcompView", "glmmTMB", "pwr", "ggh4x", "multcomp", "flextable", "car", "MASS", 'dunn.test', "dplyr","rstatix")

for (package in packages) {
  if (!requireNamespace(package, quietly = TRUE)) {
    install.packages(package)
  }
}
lapply(packages, library, character.only = TRUE)
```

## subset for this year's data

```{r}
SCP<- SCP[SCP$site == "Westport",]
LLTemp<- LLTemp[LLTemp$site == "CT",] 
LLTime<- LLTime[LLTime$site == "CT",] 

#this is subsetting to only include the city they came from this year
#I called it CT or Westport in diff datasets
```

## how many eggs (total over both shipments) were used to measure SCPs? 

```{r}
#by collection
SCP %>%
  group_by(collection_date) %>%
  summarise(n = n()) 

#total
nrow(SCP)
```

## What was the min, median, and max SCP?  

```{r}
SCP %>%
  summarize(min = min(SCP), median = median(SCP), max=max(SCP))
```

## How many egg masses and eggs were tested for Lower Lethal Temperature studies (again over both collections)?  

```{r}
#by collection
LLTemp %>%
  group_by(collection_date) %>%
  summarise(n = n()) 

#total
nrow(LLTemp)
```

### number of egg *masses*: 
we used 6 egg masses * 3 runs 

= 18 egg masses /shipment

= 36 egg masses total

## What temperature seemed to be responsible for 50% mortality (based on BVT)? 

```{r}
#remove dried eggs
LLTemp <- LLTemp[is.na(LLTemp$remove),] 

#remove controls, since these were only used to calculate viability, done in Excel
LLTemp <- LLTemp[!(LLTemp$treatment %in% c("Live Control", "Dead Control")), ] 

#add right and left bounds of interval censoring
LLTemp <- LLTemp %>% 
  mutate(left = ifelse(viability == 1, temp, Inf),
         right = ifelse(viability == 1, -Inf, temp)) 
#if live, left = removal temp, right = -infinity
#if dead, left = infinity, right = removal temp

#generate Kaplan-Meier curves
surv_obj <- with(LLTemp, Surv(time = right, time2 = left, type = "interval2"))
km_fit <- survfit(surv_obj ~ month + run, data = LLTemp)

#exract data
survd <- data.frame(temp = km_fit$time,
                        propalive = km_fit$surv, 
                        lower = km_fit$lower, 
                        upper = km_fit$upper)
survd_newdata <- data.frame(strata = km_fit$strata)
survd_newdata <- survd_newdata %>%
  rownames_to_column(var = "var") %>%
  separate(col = "var", 
           into = c("month", "run"),
           sep = ",") %>%
  mutate(month = sub(".*=", "", month)) %>%
  mutate(run = sub(".*=", "", run)) %>%
  slice(rep(row_number(), strata))
survd <- cbind(survd, survd_newdata[,-3]) %>% #-3 is removing useless third column
  mutate(
    month = str_replace_all(month, "\\s+", ""),
    run = str_replace_all(run, "\\s+", ""),
  propalive = 1-propalive, #take the inverse, as to display survival not mortality
  lower = 1-lower, #take the inverse
  upper = 1-upper) #take the inverse

#fit model, here im keeping sitemonth in it just because
survd <- survd %>% mutate(propalive = ifelse(propalive == 0, 0.001, ifelse(propalive == 1, 0.999, survd$propalive)))
LLTempModel <- glmmTMB(propalive ~ temp * month + (1|run) -1 -temp, data = survd, family = beta_family(link = "logit"))
summary(glht(LLTempModel), adjust = "bonferroni")
#shows intercept and slope for both months

#make predictions of LTs
#make newdata
newdata <- expand.grid(month = unique(survd$month), temp = seq(10, -150, by = -0.1))

#to acheieve confidence interval between 0 and 1, 
#calculate fit and SE with predict on the LINK function (to get logit link values), 
#then back tranform them to true values
LLTempP <- cbind(newdata,
                 predict(LLTempModel, newdata, type = "link", se.fit = T, re.form = NA))
LLTempP$lower <- LLTempP$fit - 1.96 * LLTempP$se.fit
LLTempP$upper <- LLTempP$fit + 1.96 * LLTempP$se.fit
LLTempP$fit <- plogis(LLTempP$fit)
LLTempP$upper <- plogis(LLTempP$upper)
LLTempP$lower <- plogis(LLTempP$lower)

#create function for inverse predictions
calculate_quantiles <- function(data) {
  list(
    data.frame(
      quantile = "50%", 
      estimate = data[which.min(abs(data$fit - 0.5)), ]$temp,
      lower = data[which.min(abs(data$lower - 0.5)), ]$temp, 
      upper = data[which.min(abs(data$upper - 0.5)), ]$temp
    ),
    data.frame(
      quantile = "90%", 
      estimate = data[which.min(abs(data$fit - 0.1)), ]$temp,
      lower = data[which.min(abs(data$lower - 0.1)), ]$temp, 
      upper = data[which.min(abs(data$upper - 0.1)), ]$temp
    )
  )
}

#retrieve quantiles
quantiles <- LLTempP %>%
  group_by(month) %>%
  do(bind_rows(calculate_quantiles(.))) %>%
  ungroup()

#these are the LT50 and 90s with 95% confidence intervals
quantiles
#there are really large errors for the January data but eventually they'd be combined

```

## Lastly, how many egg masses (and eggs) were used in Lower Lethal Time experiments?

```{r}
#by collection
LLTime %>%
  group_by(collection_date) %>%
  summarise(n = n()) 

#total
nrow(LLTime)
```

### number of egg *masses*: 
we used 5 egg masses in +3, 7 in -18, 7 in -22 

= 19 egg masses /shipment

= 38 egg masses total
 
 
## LTs for LLTime, couldn't get these

```{r}
# #remove dried eggs
# LLTime <- LLTime[is.na(LLTime$remove),] 
# 
# #make time a category
# LLTime$time <- as.factor(LLTime$time)
# 
# #add right and left bounds of interval censoring
# LLTime <- LLTime %>% 
#   mutate(left = ifelse(viability == 1, time, Inf),
#          right = ifelse(viability == 1, -Inf, time)) 
# #if live, left = removal time, right = -infinity
# #if dead, left = infinity, right = removal time
# 
# #generate Kaplan-Meier curves
# surv_obj <- with(LLTime, Surv(time = right, time2 = left, type = "interval2"))
# km_fit <- survfit(surv_obj ~ temp + collection_date, data = LLTime)
# km_fit ##can see LT50s by run
# 
# #exract data
# survd <- data.frame(time = km_fit$time,
#                         propalive = km_fit$surv, 
#                         lower = km_fit$lower, 
#                         upper = km_fit$upper)
# survd_newdata <- data.frame(strata = km_fit$strata)
# survd_newdata <- survd_newdata %>%
#   rownames_to_column(var = "var") %>%
#   separate(col = "var", 
#            into = c("temp", "collection_date"),
#            sep = ",") %>%
#   mutate(temp = sub(".*=", "", temp)) %>%
#   mutate(run = sub(".*=", "", collection_date)) %>%
#   slice(rep(row_number(), strata))
# survd <- cbind(survd, survd_newdata[,-3]) %>% #-3 is removing useless third column
#   mutate(
#     temp = str_replace_all(temp, "\\s+", ""),
#     run = str_replace_all(collection_date, "\\s+", ""))
# 
# #fit model, here im using month as a random effect
# #i think we should be using collection date as a random effect when it's not significant
# survd <- survd %>% mutate(propalive = ifelse(propalive == 0, 0.001, ifelse(propalive == 1, 0.999, survd$propalive)))
# LLTimeModel <- glmmTMB(propalive ~ time * temp + (1|collection_date) -1 -time, data = survd, family = beta_family(link = "logit"))
# summary(glht(LLTimeModel), adjust = "bonferroni")
# #shows intercept and slope for both temps
# 
# #make predictions of LTs
# #make newdata
# newdata <- expand.grid(temp = unique(survd$temp), time = seq(0, 10, by = 0.1))
# 
# #to acheieve confidence interval between 0 and 1, 
# #calculate fit and SE with predict on the LINK function (to get logit link values), 
# #then back tranform them to true values
# LLTimeP <- cbind(newdata,
#                  predict(LLTimeModel, newdata, type = "link", se.fit = T, re.form = NA))
# LLTimeP$lower <- LLTimeP$fit - 1.96 * LLTimeP$se.fit
# LLTimeP$upper <- LLTimeP$fit + 1.96 * LLTimeP$se.fit
# LLTimeP$fit <- plogis(LLTimeP$fit)
# LLTimeP$upper <- plogis(LLTimeP$upper)
# LLTimeP$lower <- plogis(LLTimeP$lower)
# 
# ggplot()+
#   geom_point(data = survd, aes(x = time, y = propalive, color = temp, shape = temp))+
#   geom_errorbar(data = survd, aes(x = time, ymin = lower, ymax = upper, color = temp), alpha = 0.8) +
#   geom_line(data = LLTimeP, aes(x = time, y = fit, color = temp))+
#   geom_ribbon(data = LLTimeP, aes(x = time, ymin = lower, ymax = upper, fill = temp), alpha = .1)
# 
# #create function for inverse predictions
# calculate_quantiles <- function(data) {
#   list(
#     data.frame(
#       quantile = "50%", 
#       estimate = data[which.min(abs(data$fit - 0.5)), ]$time,
#       lower = data[which.min(abs(data$lower - 0.5)), ]$time, 
#       upper = data[which.min(abs(data$upper - 0.5)), ]$time
#     ),
#     data.frame(
#       quantile = "90%", 
#       estimate = data[which.min(abs(data$fit - 0.9)), ]$time,
#       lower = data[which.min(abs(data$lower - 0.9)), ]$time, 
#       upper = data[which.min(abs(data$upper - 0.9)), ]$time
#     )
#   )
# }
# 
# #retrieve quantiles
# quantiles <- LLTimeP %>%
#   group_by(temp) %>%
#   do(bind_rows(calculate_quantiles(.))) %>%
#   ungroup()
# 
# #these are the LT50 and 90s with 95% confidence intervals
# quantiles


```
